---
title: "pmlRaw"
author: "Calvin Seto"
date: "January 4, 2016"
output: html_document
---

# Data Processing
The test set has 20 observations of 160 variables.  Use sapply and a function using all and is.na to test data frame, pmlTest0 and find any column that contained only missing values.
The following algorithm will select variables with data for every observation in the data set.
Create a vector of missing value counts in the data frame, pmlTest0 and use it to create a vector of missing value count = 0, i.e. vector of pmlTest0 variables with no missing values, which is used to create the subset pmlTest1.
```{r cache=TRUE}
setwd("~/Dropbox/jhudatascience/8_Practical_Machine_Learning/CourseProject")
pmlTest0 <- read.csv("data/pml-testing.csv")
pmlTest0MissingCounts <- sapply(pmlTest0, function(x)sum(is.na(x)))
colNamesTest <- pmlTest0MissingCounts[pmlTest0MissingCounts==0]
pmlTest1 <- pmlTest0[,names(colNamesTest)]
```

The training set has 19,622 observations of 160 variables.  Numeric variables are coming over as factors.  There are missing values and Excel divide by zero errors.
Use stringsAsFactors=FALSE and na.strings = c("#DIV/0!","","NA") to fix numeric variables and convert #DIV/0!, "", and NA to NA.
```{r cache=TRUE}
setwd("~/Dropbox/jhudatascience/8_Practical_Machine_Learning/CourseProject")
pmlTrain0 <- read.csv("data/pml-training.csv")
str(pmlTrain0)
sum(is.na(pmlTrain0)) # 1,287,472 the DIV/0! values are here
pmlTrain1 <- read.csv("data/pml-training.csv", stringsAsFactors = FALSE, na.strings = c("#DIV/0!","","NA"))
sum(is.na(pmlTrain1)) # 1,925,102 the DIV/0! values are converted to NA properly
```

Use sapply and a function using all and is.na to test data frame, pmlTrain1 and find any column that contained only missing values.  This resulted in 6 columns: kurtosis_yaw_belt, skewness_yaw_belt, kurtosis_yaw_dumbbell, skewness_yaw_dumbbell, kurtosis_yaw_forearm, and skewness_yaw_forearm.
```{r cache=TRUE}
sapply(pmlTrain1, function(x)all(is.na(x))) #6 variables contain all NA
sum(is.na(pmlTrain1[,"kurtosis_yaw_belt"])) # 19622 count to confirm
sum(is.na(pmlTrain1[,"skewness_yaw_belt"])) # 19622
sum(is.na(pmlTrain1[,"kurtosis_yaw_dumbbell"])) # 19622
sum(is.na(pmlTrain1[,"skewness_yaw_dumbbell"])) # 19622
sum(is.na(pmlTrain1[,"kurtosis_yaw_forearm"])) # 19622
sum(is.na(pmlTrain1[,"skewness_yaw_forearm"])) # 19622
```

The following algorithm will select variables with data for every observation in the data set.
Create a vector of missing value counts in the data frame, pmlTrain1 and use it to create a vector of missing value count = 0, i.e. vector of pmlTrain1 variables with no missing values, which is used to create the subset pmlTrain2.
```{r cache=TRUE}
pmlTrain1MissingCounts <- sapply(pmlTrain1, function(x)sum(is.na(x)))
pmlTrain1Complete <- pmlTrain1MissingCounts[pmlTrain1MissingCounts==0]
pmlTrain2 <- pmlTrain1[,names(pmlTrain1Complete)]
```

This code verifies that the processed training and testing data sets are the same except for the problem_id variable in the testing set and the classe variable in the training set.  We've reduced the 160 variables to 60 variables. 
```{r cache=TRUE}
setequal(names(pmlTest1),names(pmlTrain2))
setdiff(names(pmlTest1),names(pmlTrain2))
setdiff(names(pmlTrain2),names(pmlTest1))
intersect(names(pmlTest1),names(pmlTrain2))
```

# Exploratory Data Analysis
1"X" (observation ID)
2"user_name" "carlitos" "pedro"    "adelmo"   "charles"  "eurico"   "jeremy"
3"raw_timestamp_part_1" 1322832937
4"raw_timestamp_part_2" 204334
5"cvtd_timestamp" 02/12/2011 13:35
sort(unique(pmlTrain2$cvtd_timestamp))
"02/12/2011 13:32" "02/12/2011 13:33" "02/12/2011 13:34" "02/12/2011 13:35" 
"02/12/2011 14:56" "02/12/2011 14:57" "02/12/2011 14:58" "02/12/2011 14:59"
"05/12/2011 11:23" "05/12/2011 11:24" "05/12/2011 11:25"
"05/12/2011 14:22" "05/12/2011 14:23" "05/12/2011 14:24"
"28/11/2011 14:13" "28/11/2011 14:14" "28/11/2011 14:15"
"30/11/2011 17:10" "30/11/2011 17:11" "30/11/2011 17:12"
-4 days (Dec 2 and 5 of 2011 Nov 28 and 30 of 2011)
-6 attempts
Dec 2 and 5 of 2011 13:32 to 13:35, 14:56 to 14:59, 11:23 to 11:25, 14:22 to 14:24
Nov 28 and 30 of 2011 14:13 to 14:15, 17:10 to 17:12

6/new_window" no yes
7"num_window" 1-864           
A sliding window approach was used for feature extraction.
Different lengths from 0.5 sec to 2.5 sec, with 0.5 sec overlap
Resulting in total 96 derived feature sets.

Sliding Windows are an approach to the sequential supervised learning problem. mlsd-ssspr.pdf

There are 13 variables for belt, arm, dumbbell, and forearm. 13x4=52.
names(pmlTrain2[8:20])
names(pmlTrain2[21:33])
names(pmlTrain2[34:46])
names(pmlTrain2[47:59])

"roll_belt"           
"pitch_belt"           "yaw_belt"             "total_accel_belt"     "gyros_belt_x"        
"gyros_belt_y"         "gyros_belt_z"         "accel_belt_x"         "accel_belt_y"        
"accel_belt_z"         "magnet_belt_x"        "magnet_belt_y"        "magnet_belt_z"       

"roll_arm"             "pitch_arm"            "yaw_arm"              "total_accel_arm"     
"gyros_arm_x"          "gyros_arm_y"          "gyros_arm_z"          "accel_arm_x"         
"accel_arm_y"          "accel_arm_z"          "magnet_arm_x"         "magnet_arm_y"        
"magnet_arm_z"

"roll_dumbbell"        "pitch_dumbbell"       "yaw_dumbbell"        
"total_accel_dumbbell" "gyros_dumbbell_x"     "gyros_dumbbell_y"     "gyros_dumbbell_z"    
"accel_dumbbell_x"     "accel_dumbbell_y"     "accel_dumbbell_z"     "magnet_dumbbell_x"   
"magnet_dumbbell_y"    "magnet_dumbbell_z"

"roll_forearm"         "pitch_forearm"       
"yaw_forearm"          "total_accel_forearm"  "gyros_forearm_x"      "gyros_forearm_y"     
"gyros_forearm_z"      "accel_forearm_x"      "accel_forearm_y"      "accel_forearm_z"     
"magnet_forearm_x"     "magnet_forearm_y"     "magnet_forearm_z"

60 "classe" or "problem_id"

# Data Splitting
```{r cache=TRUE}
library(caret);
inTrain <- createDataPartition(y=pmlTrain2$classe,
                               p=0.75, list=FALSE)
training <- pmlTrain2[inTrain,]
testing <- pmlTrain2[-inTrain,]
dim(training)
```


data exploration plotting predictors and looking for patterns, imbalance, outliers, unexplained groups, skewed variables
-summary of predictors min, max, mean, median, quartiles
-build training and testing set and explore training only and leave testing as hold out set
-featurePlot show outcome and predictors scatterplots
-qplot scatterplot of outcome and predictor (add color by variables; add regression smoothers to see trends)
-make numbers into categories or factors with cut2 plot with plot (add boxplots and overlay points)
-tables on factors and predictor showing counts or proportions prop.table
-density plots for continuous predictors to see where the bulk of the data is color by belt, arm, dumbbell, and forearm?

# table of classe in training
   A    B    C    D    E 
4185 2848 2567 2412 2706 

```{r cache=TRUE}
summary(as.factor(training$classe))
```

# scatterplot of important variables colored by classe
```{r}
qplot(x=roll_belt, y=yaw_belt,data=training,colour=classe)
qplot(x=roll_belt, y=pitch_belt,data=training,colour=classe)
qplot(x=pitch_belt, y=yaw_belt,data=training,colour=classe)
qplot(x=roll_belt, y=magnet_dumbbell_z,data=training,colour=classe)
qplot(x=roll_belt, y=pitch_forearm,data=training,colour=classe)
```

# density plot belt predictors by classe
```{r cache=TRUE}
qplot(roll_belt,colour=classe,data=training,geom="density")
qplot(pitch_belt,colour=classe,data=training,geom="density")
qplot(yaw_belt,colour=classe,data=training,geom="density")
qplot(total_accel_belt,colour=classe,data=training,geom="density")
qplot(gyros_belt_x,colour=classe,data=training,geom="density")
qplot(gyros_belt_y,colour=classe,data=training,geom="density")
qplot(gyros_belt_z,colour=classe,data=training,geom="density")
qplot(accel_belt_x,colour=classe,data=training,geom="density")
qplot(accel_belt_y,colour=classe,data=training,geom="density")
qplot(accel_belt_z,colour=classe,data=training,geom="density")
qplot(magnet_belt_x,colour=classe,data=training,geom="density")
qplot(magnet_belt_y,colour=classe,data=training,geom="density")
qplot(magnet_belt_z,colour=classe,data=training,geom="density")
```

# density plot arm predictors by classe
```{r cache=TRUE}
qplot(roll_arm,colour=classe,data=training,geom="density")
qplot(pitch_arm,colour=classe,data=training,geom="density")
qplot(yaw_arm,colour=classe,data=training,geom="density")
qplot(total_accel_arm,colour=classe,data=training,geom="density")
qplot(gyros_arm_x,colour=classe,data=training,geom="density")
qplot(gyros_arm_y,colour=classe,data=training,geom="density")
qplot(gyros_arm_z,colour=classe,data=training,geom="density")
qplot(accel_arm_x,colour=classe,data=training,geom="density")
qplot(accel_arm_y,colour=classe,data=training,geom="density")
qplot(accel_arm_z,colour=classe,data=training,geom="density")
qplot(magnet_arm_x,colour=classe,data=training,geom="density")
qplot(magnet_arm_y,colour=classe,data=training,geom="density")
qplot(magnet_arm_z,colour=classe,data=training,geom="density")
```

# density plot dumbbell predictors by classe
```{r cache=TRUE}
qplot(roll_dumbbell,colour=classe,data=training,geom="density")
qplot(pitch_dumbbell,colour=classe,data=training,geom="density")
qplot(yaw_dumbbell,colour=classe,data=training,geom="density")
qplot(total_accel_dumbbell,colour=classe,data=training,geom="density")
qplot(gyros_dumbbell_x,colour=classe,data=training,geom="density")
qplot(gyros_dumbbell_y,colour=classe,data=training,geom="density")
qplot(gyros_dumbbell_z,colour=classe,data=training,geom="density")
qplot(accel_dumbbell_x,colour=classe,data=training,geom="density")
qplot(accel_dumbbell_y,colour=classe,data=training,geom="density")
qplot(accel_dumbbell_z,colour=classe,data=training,geom="density")
qplot(magnet_dumbbell_x,colour=classe,data=training,geom="density")
qplot(magnet_dumbbell_y,colour=classe,data=training,geom="density")
qplot(magnet_dumbbell_z,colour=classe,data=training,geom="density")
```

# density plot forearm predictors by classe
```{r cache=TRUE}
qplot(roll_forearm,colour=classe,data=training,geom="density")
qplot(pitch_forearm,colour=classe,data=training,geom="density")
qplot(yaw_forearm,colour=classe,data=training,geom="density")
qplot(total_accel_forearm,colour=classe,data=training,geom="density")
qplot(gyros_forearm_x,colour=classe,data=training,geom="density")
qplot(gyros_forearm_y,colour=classe,data=training,geom="density")
qplot(gyros_forearm_z,colour=classe,data=training,geom="density")
qplot(accel_forearm_x,colour=classe,data=training,geom="density")
qplot(accel_forearm_y,colour=classe,data=training,geom="density")
qplot(accel_forearm_z,colour=classe,data=training,geom="density")
qplot(magnet_forearm_x,colour=classe,data=training,geom="density")
qplot(magnet_forearm_y,colour=classe,data=training,geom="density")
qplot(magnet_forearm_z,colour=classe,data=training,geom="density")
```

## histgrams - belt
```{r cache=TRUE}
with(training, hist(roll_belt))
with(training, hist(pitch_belt))
with(training, hist(yaw_belt))
with(training, hist(total_accel_belt))
with(training, hist(gyros_belt_x))
with(training, hist(gyros_belt_y))
with(training, hist(gyros_belt_z))
with(training, hist(accel_belt_x))
with(training, hist(accel_belt_y))
with(training, hist(accel_belt_z))
with(training, hist(magnet_belt_x))
with(training, hist(magnet_belt_y))
with(training, hist(magnet_belt_z))
```

## histgrams - arm
```{r cache=TRUE}
with(training, hist(roll_arm))
with(training, hist(pitch_arm))
with(training, hist(yaw_arm))
with(training, hist(total_accel_arm))
with(training, hist(gyros_arm_x))
with(training, hist(gyros_arm_y))
with(training, hist(gyros_arm_z))
with(training, hist(accel_arm_x))
with(training, hist(accel_arm_y))
with(training, hist(accel_arm_z))
with(training, hist(magnet_arm_x))
with(training, hist(magnet_arm_y))
with(training, hist(magnet_arm_z))
```

## histgrams - dumbbell
```{r cache=TRUE}
with(training, hist(roll_dumbbell))
with(training, hist(pitch_dumbbell))
with(training, hist(yaw_dumbbell))
with(training, hist(total_accel_dumbbell))
with(training, hist(gyros_dumbbell_x))
with(training, hist(gyros_dumbbell_y))
with(training, hist(gyros_dumbbell_z))
with(training, hist(accel_dumbbell_x))
with(training, hist(accel_dumbbell_y))
with(training, hist(accel_dumbbell_z))
with(training, hist(magnet_dumbbell_x))
with(training, hist(magnet_dumbbell_y))
with(training, hist(magnet_dumbbell_z))
```

## histgrams - forearm
```{r cache=TRUE}
with(training, hist(roll_forearm))
with(training, hist(pitch_forearm))
with(training, hist(yaw_forearm))
with(training, hist(total_accel_forearm))
with(training, hist(gyros_forearm_x))
with(training, hist(gyros_forearm_y))
with(training, hist(gyros_forearm_z))
with(training, hist(accel_forearm_x))
with(training, hist(accel_forearm_y))
with(training, hist(accel_forearm_z))
with(training, hist(magnet_forearm_x))
with(training, hist(magnet_forearm_y))
with(training, hist(magnet_forearm_z))
```

# summary - belt, arm, dumbbell, forearm
```{r cache=TRUE}
summary(training[8:20])
summary(training[21:33])
summary(training[34:46])
summary(training[47:59])
```

# caret featurePlots - belt
```{r cache=TRUE}
featurePlot(x=training[,c("roll_belt","pitch_belt","yaw_belt")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("total_accel_belt","accel_belt_x","accel_belt_y","accel_belt_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("gyros_belt_x","gyros_belt_y","gyros_belt_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("magnet_belt_x","magnet_belt_y","magnet_belt_z")],
            y=training$classe,
            plot="pairs")
```


# caret featurePlots - arm
```{r cache=TRUE}
featurePlot(x=training[,c("roll_arm","pitch_arm","yaw_arm")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("total_accel_arm","accel_arm_x","accel_arm_y","accel_arm_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("gyros_arm_x","gyros_arm_y","gyros_arm_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("magnet_arm_x","magnet_arm_y","magnet_arm_z")],
            y=training$classe,
            plot="pairs")
```

# caret featurePlots - dumbbell
```{r cache=TRUE}
featurePlot(x=training[,c("roll_dumbbell","pitch_dumbbell","yaw_dumbbell")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("total_accel_dumbbell","accel_dumbbell_x","accel_dumbbell_y","accel_dumbbell_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("magnet_dumbbell_x","magnet_dumbbell_y","magnet_dumbbell_z")],
            y=training$classe,
            plot="pairs")
```

# caret featurePlots - forearm
```{r cache=TRUE}
featurePlot(x=training[,c("roll_forearm","pitch_forearm","yaw_forearm")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("total_accel_forearm","accel_forearm_x","accel_forearm_y","accel_forearm_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("gyros_forearm_x","gyros_forearm_y","gyros_forearm_z")],
            y=training$classe,
            plot="pairs")
featurePlot(x=training[,c("magnet_forearm_x","magnet_forearm_y","magnet_forearm_z")],
            y=training$classe,
            plot="pairs")
```



Here are the column numbers for each of the 4 sensor locations
```{r cache=TRUE}
names(training[8:20]) # belt
names(training[21:33]) # arm
names(training[34:46]) # dumbbell
names(training[47:59]) # forearm

```

# Train options
train parameters
method
data
preProcess
weights
metric cat acc cont rmse

train control
method (resampling) boot boot632 cv repeatedcv LOOCV
number (iterations)
repeats
p size of training
initialWindow for time points
horizon for time points
savePredictions
summaryFunction
preProcOptions
predictionBounds
seeds
allowParallel

# Boosting

# Model Fit
```{r cache=TRUE}
# library(caret)
# library(doMC)
# registerDoMC(4)
# this took 6 hours!
# Sys.time()
# grid <- expand.grid(mtry=100)
# fitControl <- trainControl(## 10-fold CV
                           # method = "repeatedcv",
                           # number = 10,
                           # repeated ten times
                           # repeats = 10)

# modFit <- train(x=pmlTrain2[,8:59],y=pmlTrain2[,60],
                # method="rf",
                # prox=TRUE,
                # tuneGrid = grid,
                # trControl = fitControl)
# Sys.time()
# modFit
```


predictions <- predict(fitRF, newdata=pmlTest1)

